---
title: "Asger_Krogh_2021"
author: "Asger Krogh"
date: "04 01 2022"
output: html_document
---

```{r setup, echo = FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Packages needed

```{r Find the correct custom package, include=FALSE}
#if(!require("devtools")) install.packages("devtools")
#devtools::install_github("sebastianbarfort/mapDK")
#devtools::install_github("56north/leafletDK") # install devtools if needed
```

```{r install and load packages, include=TRUE, echo = FALSE}
library(ggplot2)
library(mapDK)
library(tidyverse)
library(gganimate)
library(leafletDK)
library(leaflet.extras)
library(readr)
library(knitr)
library(stringi)
```


# Loading of Data.
The data had been downloaded with an API from "Danmarkstatistik" into OpenRefine and cleaned. 
I will the rest of the modifying here in RStudio.

I load the data from the year 2007, of number of people, aged 18, who have moved to Copenhagen. This is then specifies in which municipality they come from.
Then I check the first 6 rows to confirm it looks correct and to see the specifications of the colloum.

```{r load data fra årene 2006-2020}
data07 <- read_csv("aar2007.csv",show_col_types = FALSE)

head(data07)
```

I here want to see how big a percentage of the combined municipality population they people who moves away constitute.
I again use an API from Danmarkstatistik to find the total population of 18 year old from each municipality.

```{r load new API and mutate the colloum in here}
aar1808 <- read_csv("antal18.csv", show_col_types = FALSE)
head(aar1808)
```

# Cleaning data to use

I merge the two different datasets with the mutate function, as we can see below, the data07 set now has 5 variables
```{r mutate from data07 to aar1808}
data07 %>% 
  mutate(Total18=aar1808$INDHOLD) -> data07

head(data07)
```


```{r decimille størrelse}
data07 %>% 
  mutate(procent = (INDHOLD/Total18)*100) -> data07

head(data07)
#The new colloum i add to the existing dataset since i want to keep working with it.
```
I use the "mutate" function because i wan't to use data that already have included in my sheet. I use
the mutate to make a new colloum that shows the percentage of the population that moves away. 

Since i use a custom and very specific packages, the minicipalityDK and mapDK, they cannot read my data since the municipality names must match up 100%. I already cleaned up the special charracters in OpenRefine, but now i need to rename the colloums as well, and make them into lowercase.
```{r omdøb til lille startbogstav}
#for 2007
data07 %>% 
  mutate(FRAKOMMUNE = tolower(FRAKOMMUNE)) -> data07

data07 %>% 
  rename(kommune = FRAKOMMUNE) -> data07

head(data07)
```

# Maps and graphs

I here use mapDK to create a graph that shows the promille who moved. This is visuallied with dark blue as low values, and lighter is higher values. 

## kommunekort1 - First municipality map 
```{r mapDK 2006}
kommunekort1 <- mapDK(values = 'procent', id = 'kommune', data = data07)

kommunekort1
```

Since i don't have the data for christiansoe and taarnby i will get a warning since they can't be included then. Also i have no value for Copenhagen, since you can't move from and too the same. 

I find that this isn't very easy to read, no i will now try with the municipalityDK. 

This way it becomes a little easier to read, and then the map is also interactive now, you can click on the municipalities and see the given value.

## kommunekort2 - New municipality map 
```{r municipalityDK}
kommunekort2 <- municipalityDK("procent", "kommune", data = data07, legend=T,pal = "GnBu") %>%
  setMapWidgetStyle(list(background= "white"))

kommunekort2
```


I wan't to make the contrast of colours even more clear and the map easier to use and understand. 
```{r farver klar, echo = FALSE}
library(RColorBrewer)

colfunc <- colorRampPalette(c("#2C77BF", "red"))
```


## kommunekort3 - First municipality map with colours
```{r med nye farver}
kommunekort3 <- municipalityDK("procent", "kommune", data = data07, legend=T,pal = colfunc(10)) %>%
  setMapWidgetStyle(list(background= "white"))

kommunekort3
```

The contrast now is from blue to red, and therfor much easier to undersand and read. 


# Graphs for Frederiksberg and Lolland via API
I see that Frederiksberg i the municipality with the highest values. 
Now i want to examine if that has changed over time.


I use the Danmarkstatistik API again and load the data:

```{r frederiksbergapi}
fre18 <- read_delim("https://api.statbank.dk/v1/data/FLY66/CSV?delimiter=Semicolon&TILKOMMUNE=101&FRAKOMMUNE=147&ALDER=18&Tid=*",show_col_types = FALSE)
```

I then want to do a simple plot with ggplot to see if this is normal values or an outlier.
```{r  frederiksbergplot}
ggplot(fre18) +  aes(x = TID, y = INDHOLD, colour = "red") + geom_path()
```

I can here see that the year 2008, isn't an outlier and the fact that for Frederiksberg is the most popuplar place to move is Copenhagen, seems very likely. 


Then to test the relation between Copenhagen and the souther part of Sjælland:
```{r frederiksberg}
lol18 <- read_delim("https://api.statbank.dk/v1/data/FLY66/CSV?delimiter=Semicolon&TILKOMMUNE=101&FRAKOMMUNE=360&ALDER=18&Tid=*",show_col_types = FALSE)
```

I then want to do a simple plot with ggplot to see if this is normal values or an outlier.
```{r plot lolland}
ggplot(lol18) +  aes(x = TID, y = INDHOLD, colour = "red") + geom_path()
```

The graph shows a very simular development to the Frederiksberg graph, and indicates a decline up towards 2020. 


# Data from 2017 and maps to comparason
Finally i want to test that my project is reproducable and i also wont to make my kommunekort3 comparabel with another visualization. 
Therefor i want to make the same map, but with data for 2017, 10 years later, to see if there are changes. 


```{r load data fra år 2017 antal}
aar17 <- read_csv("aar1817ny.csv",show_col_types = FALSE)
```

```{r kommunedata fra 2017}
data17 <- read_csv("Data17.csv",show_col_types = FALSE)
```


```{r from  data07 to aar1808 complete 2017, include=TRUE, eval=TRUE}
data17 %>% 
  mutate(Total18=aar17$INDHOLD) -> data17
```

```{r omdøb til lille startbogstav 2017, include=TRUE, eval=TRUE}
#for 2017
data17 %>% 
  mutate(FRAKOMMUNE= tolower(FRAKOMMUNE)) -> data17

data17 %>% 
  rename(kommuner = FRAKOMMUNE ) -> data17
```


```{r decimille størrelse 2017, include=TRUE, eval=TRUE}
data17 %>% 
  mutate(procent = (INDHOLD/Total18)*100) -> data17
#The new colloum i add to the existing dataset since i want to keep working with it.
```

```{r for 2017 igen, include=TRUE, eval=TRUE}
kommunekort4 <- municipalityDK("procent", "kommuner", data = data17, legend=T,pal = colfunc(10)) %>%
  setMapWidgetStyle(list(background= "white"))
```

I can easily reproduce my project with new data.

```{r run last map 2017, include=TRUE, eval=TRUE}
kommunekort4
```
And for comparrison with my original map for 2007.

```{r run last map 2007, include=TRUE, eval=TRUE}
kommunekort3
```